---
title: "Godlike"
author: "Petra ([oleumpetra@gmail.com](mailto:oleumpetra@gmail.com))"
---

```{r initialsetup, include=FALSE}
library(knitr)
library(broom)
library(ggplot2)
library(scales)
opts_chunk$set(tidy.opts=list(width.cutoff=30), tidy=TRUE)
options(width=60, knitr.kable.NA = '—')
set.seed(12345)

roll10s <- function(dicepool, hard=0, h.value=7) {
    c(sample(1:10, dicepool, replace=TRUE), rep(h.value, hard))
}

dable <- function(x) {
    tx <- table(x)
    dx <- broom::tidy(tx)

    names(dx) <- c("Number on die", "Frequency")
    kable(dx, caption=paste0("Dice roll: ", length(x), "d10"))

}

maxwidth <- function(x) {
    max(table(x))
}
 
maxmatchheight <- function(x) {
    tx <- table(x)
    tx <- subset(tx, tx >= 2)
    if (length(tx) == 0) {
        return(NA)
    }
    return(as.numeric(names(tx)[length(tx)]))
}

nummatches <- function(x) {
    tx <- table(x)
    tx <- subset(tx, tx >= 2)
    length(tx)
}

maxheight <- function(x) {
    tx <- table(x)
    mw <- maxwidth(x)
    tx <- subset(tx, tx == mw)
    return(as.numeric(names(tx)[length(tx)]))
}

manyrolls <- function(dicepool, replicates=2500) {
    blank=rep(NA, replicates)
    results <- list("Width"=blank, "Height"=blank, "Matches"=blank, "Best.Height"=blank)
    for (i in 1:replicates) {
        throw <- roll10s(dicepool)
        results$Width[i] = maxwidth(throw)
        results$Height[i] = maxheight(throw)
        results$Matches[i] = nummatches(throw)
        results$Best.Height[i] = maxmatchheight(throw)
    }
    results <- as.data.frame(results)
    names(results) <- c("Maximum width", "Height of maximum width", "Number of matches", "Best overall match height")
    return(results)
}

proportions <- function(tx) {
    dx <- broom::tidy(table(tx))
    names(dx) <- c("Result", "Frequency")
    dx <- cbind(dx, Percentage=paste0(round(dx$Frequency/length(tx) * 100, 2), "%"))
    return(dx)
}
 
proptable.maxwidth <- function(dx, dicepool) {
    prt <- proportions(dx[,"Maximum width"])
    names(prt)[1] <- "Width"
    kable(prt, caption=paste0("Maximum width of ", nrow(dx), " rolls of ", dicepool, "d10"))
}

dstats <- function(x) {
    tx <- table(x)
    total <- length(x)
    dx <- data.frame(Statistic=c("Maximum Width", "Height of maximum width roll", "Total number of matches", "Maximum height out of all matches"), Value=c(maxwidth(x), maxheight(x), nummatches(x), maxmatchheight(x)))
    kable(dx, caption=paste0("Statistics for ", length(x), "d10 dice roll"))
}
```

_Godlike_ operates on a system based around `d10`s, where success is determined by rolling "matches" of particular "depth."

```{r rolling, include=FALSE}
five.d10 <- roll10s(5)
```

For example, a roll of `5d10` might produce the numbers `r five.d10`. Or, in table format:

```{r dicetable, echo=FALSE}
dable(five.d10)
```

We have two eights and two nines---both "matches"! Each of these matches has a "width" of 2, but the pair of nines has a "height" of nine, while the eights only have a height of eight. 

```{r dice2, echo=FALSE}
dstats(five.d10)
```

You generally need a match of _some_ kind to succeed, but sometimes height is more important than weight, and sometimes it is the other way around. In the context of "dynamic contests" the book says:

> "Whoever rolls higher (see Height) wins the contest, while whoever rolls wider (see Width) finishes first. Running a race against others is a dynamic contest."

Meanwhile in combat, width matters for both initiative and initiative.

Regardless of the kind of roll, the way to be better is to have more dice to roll. More dice mean wider matches, and more reliable ones; it also means _more_ matches, which helps increase the chances of higher matches too. The question then is: how many dice do you need?

## Increasing height

```{r roll2, echo=FALSE}
ten.d10 <- roll10s(10)
```

Our `5d10` roll only produced two matches of width 2---which is actually pretty good. With say `10d10` (the maximum dice pool) we might be able to do better: in this case (`r ten.d10`).

```{r stats2, echo=FALSE}
dable(ten.d10)
dstats(ten.d10)
```

For this particular roll we did manage to double our matches, but we didn't improve on the width or height. Lets roll again:

```{r roll3, echo=FALSE}
ten.d10 <- roll10s(10)
dable(ten.d10)
dstats(ten.d10)
```

Now we have a set of three fives (if we want to maximise width) _and_ a pair of tens (if height is more important).

Of course, looking at individual rolls isn't actually all that useful except as examples.

```{r roll4, echo=FALSE}
twenty.5d10 <- manyrolls(5, replicates=20)
kable(twenty.5d10, caption="20 rolls of 5d10")
```

With lots of rolls we can look at a table of proportions:

```{r proptable1, echo=FALSE}
proptable.maxwidth(twenty.5d10, 5)
```

This suggests that, with `5d10`, you're probably going to get (at least one) pair, but are unlikely to get a triple. But 20 rolls isn't really a lot, so lets do it 2500 times just to make sure:

```{r roll5, echo=FALSE}
lots.5d10 <- manyrolls(5, replicates=2500)
proptable.maxwidth(lots.5d10, 5)
```

Now we see that actually the chances of a pair isn't quite so high, and that a match of width 4 is in fact possible (if rare). Theoretically we know that all five dice could in fact come up with the same number, but the chance of this occurring is $1:10^4 = 1:10,000$, so we wont hold our breath or be too mad it did not happen in this case.

Lets make a graph---this will make it easier to compare between dice pools later.

```{r graph1, echo=FALSE}
lg <- ggplot(lots.5d10, aes(x=`Maximum width`))
lg <- lg + geom_histogram(aes(y=(..count..)/sum(..count..),fill=as.factor(`Height of maximum width`)), binwidth=1)
lg <- lg + scale_y_continuous(labels=percent)
lg <- lg + labs(fill="Height", y="Percentage", title="2500 rolls of 5d10")
lg + theme_classic()
```

I've coloured the bars by the height of the rolls to illustrate an important point: if you roll really poorly in terms of matches, the numbers you do get will be relatively good. For example if you get no matches on `5d10` the worst roll you could have got is some variant of (`r 1:5`), meaning that you must have got at _least_ a 5 (probably a lot better). And if your opponent in a contested roll also didn't get any matches, and your GM is nice, that _might_ just be enough to save you.

Better yet, this same effect works with pairs too, if your dice pool is large enough. Lets roll `10d10`:

```{r roll6, echo=FALSE}
lots.10d10 <- manyrolls(10, replicates=2500)
proptable.maxwidth(lots.10d10, 10)
lg10 <- ggplot(lots.10d10, aes(x=`Maximum width`))
lg10 <- lg10 + geom_histogram(aes(y=(..count..)/sum(..count..),fill=as.factor(`Height of maximum width`)), binwidth=1)
lg10 <- lg10 + scale_y_continuous(labels=percent)
lg10 <- lg10 + labs(fill="Height", y="Percentage", title="2500 rolls of 10d10")
lg10 + theme_classic()
```

As you can see here, your chances of getting a triple is greater than your chances of getting a pair on `10d10`, _but_ your chances of getting a pair of tens as your best match is as high as your chances of getting a triple nine or ten added together. The keyword here is "best" match---having a best match be two ones, say, requires that you got no other matches; having it be two tens requires only that you got no triples or above. That latter situation is therefore more likely.
